### 2023-01-25 Strydom_et_al_2023_graph_embedding_draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@2c05016c09f13e610bf21955d0b2f421bbcf7424 🚀

Graph (or Network) embedding (fig.~\ref{fig:embedding}) is a family of
machine learning techniques, whose main task is to learn a mapping
function from a discrete graph to a continuous domain (Chami et al.,
2022; \textbf{arsov\_network\_2019?}). Their main goal is to learn a low
dimensional vector representations of the graph (embeddings), such that
its key properties (\emph{e.g.} local or global structures) are retained
in the embedding space (Yan et al., 2005). The embedding space may, but
will not necessarily, have lower dimensionality than the graph.
Ecological networks are promising candidates for the routine application
of embeddings, as they tend to possess a shared structural backbone (see
\emph{e.g.} Bramon Mora et al., 2018), which hints at structural
invariants in empirical data. Assuming that these structural invariants
are common enough, they would dominate the structure of networks, and
therefore be adequately captured by the first (lower) dimensions of an
embedding, without the need to measure derived aspects of their
structure (\emph{e.g.} motifs, paths, modularity, \ldots).

@@

Graph embeddings \emph{can} serve as a dimensionality reduction method.
For example, RDPG (Strydom et al., 2022) and t-SVD {[}truncated Singular
Value Decomposition; Poisot et al. (2021){]} typically embed networks
using fewer dimensions than the original network {[}the original network
has as many dimensions as species, and as many informative dimensions as
trophically unique species; Strydom, Dalla Riva, et al. (2021){]}. But
this is not necessarily the case -- indeed, one may perform a PCA (a
special case of SVD) to project the raw data into a subspace that
improves the efficacy of t-SNE {[}t-distributed stochastic neighbor
embedding; Maaten (2009){]}. There are many dimensionality reductions
(Anowar et al., 2021) that can be applied to an embedded network should
the need for dimensionality reduction (for example for data
visualisation) arise. In brief, many graph embeddings \emph{can} serve
as dimensionality reduction steps, but not all do, neither do all
dimensionality reduction methods provide adequate graph embedding
capacities. In the next section (and fig.~\ref{fig:illustration}), we
show how the amount of dimensionality reduction can affect the quality
of the embedding.

@@

The area used to infer the new metaweb in determines the species pool
that must be used to perform this embedding. Metawebs can be constructed
by assigning interactions in a list of species within geographic
boundaries. The upside of this approach is that information at the
country level is likely to be required for biodiversity assessments, as
countries set conservation goals at the national level (Buxton et al.,
2021), and as quantitative instruments are designed to work at these
scales (\textbf{turak\_measuring\_2017?}); specific strategies are often
enacted at smaller scales, nested within a specific country (Ray et al.,
2021). But there is no guarantee that these boundaries are meaningful.
In fact, we do not have a satisfying answer to the question of ``where
does an ecological network stop?''. Recent results by Martins et al.
(2022) suggest that networks are shaped within eco-regions, with abrupt
structural transitions from an eco-region to the next. Should this trend
hold generally, this would provide an ecologically-relevant scale at
which metawebs can be downscaled and predicted. Other solutions leverage
network-area relationships to identify areas in which networks are
structurally similar (see \emph{e.g.} Fortin et al., 2021; Galiana et
al., 2022, 2018), which requires ample pre-existing information about
the network. This suggests that inferred metawebs should be further
downscaled to allow for \emph{a posteriori} analyses; Llewelyn et al.
(2022) provide compelling evidence for the fact that when a place is
rich in endemic species, even at smaller spatial scales, the transfer of
information about interactions becomes more challenging, re-emphasizing
the need to identify guidelines for when and where network predictions
can be mapped.

@@

@@

>>>>

Graph (or Network) embedding (fig.~\ref{fig:embedding}) is a family of
machine learning techniques, whose main task is to learn a mapping
function from a discrete graph to a continuous domain (Arsov \& Mirceva,
2019; Chami et al., 2022). Their main goal is to learn a low dimensional
vector representations of the graph (embeddings), such that its key
properties (\emph{e.g.} local or global structures) are retained in the
embedding space (Yan et al., 2005). The embedding space may, but will
not necessarily, have lower dimensionality than the graph. Ecological
networks are promising candidates for the routine application of
embeddings, as they tend to possess a shared structural backbone (see
\emph{e.g.} Bramon Mora et al., 2018), which hints at structural
invariants in empirical data. Assuming that these structural invariants
are common enough, they would dominate the structure of networks, and
therefore be adequately captured by the first (lower) dimensions of an
embedding, without the need to measure derived aspects of their
structure (\emph{e.g.} motifs, paths, modularity, \ldots).

@@

Graph embeddings \emph{can} serve as a dimensionality reduction method.
For example, RDPG (Strydom et al., 2022) and t-SVD {[}truncated Singular
Value Decomposition; Poisot et al. (2021){]} typically embed networks
using fewer dimensions than the original network {[}the original network
has as many dimensions as species, and as many informative dimensions as
trophically unique species; Strydom, Dalla Riva, et al. (2021){]}. But
this is not necessarily the case -- indeed, one may perform a PCA (a
special case of SVD) to project the raw data into a subspace that
improves the efficacy of t-SNE {[}t-distributed stochastic neighbor
embedding; Maaten (2009){]}. There are many dimensionality reductions
(Anowar et al., 2021) that can be applied to an embedded network should
the need for dimensionality reduction (for example for data
visualisation) arise. In brief, many graph embeddings \emph{can} serve
as dimensionality reduction steps, but not all do, neither do all
dimensionality reduction methods provide adequate graph embedding
capacities. In the next section (and fig.~\ref{fig:embedding}), we show
how the amount of dimensionality reduction can affect the quality of the
embedding.

@@

The area used to infer the new metaweb in determines the species pool
that must be used to perform this embedding. Metawebs can be constructed
by assigning interactions in a list of species within geographic
boundaries. The upside of this approach is that information at the
country level is likely to be required for biodiversity assessments, as
countries set conservation goals at the national level (Buxton et al.,
2021), and as quantitative instruments are designed to work at these
scales (Turak et al., 2017); specific strategies are often enacted at
smaller scales, nested within a specific country (Ray et al., 2021). But
there is no guarantee that these boundaries are meaningful. In fact, we
do not have a satisfying answer to the question of ``where does an
ecological network stop?''. Recent results by Martins et al. (2022)
suggest that networks are shaped within eco-regions, with abrupt
structural transitions from an eco-region to the next. Should this trend
hold generally, this would provide an ecologically-relevant scale at
which metawebs can be downscaled and predicted. Other solutions leverage
network-area relationships to identify areas in which networks are
structurally similar (see \emph{e.g.} Fortin et al., 2021; Galiana et
al., 2022, 2018), which requires ample pre-existing information about
the network. This suggests that inferred metawebs should be further
downscaled to allow for \emph{a posteriori} analyses; Llewelyn et al.
(2022) provide compelling evidence for the fact that when a place is
rich in endemic species, even at smaller spatial scales, the transfer of
information about interactions becomes more challenging, re-emphasizing
the need to identify guidelines for when and where network predictions
can be mapped.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Arsov2019Network}{}}%
Arsov, N., \& Mirceva, G. (2019). \emph{Network Embedding: An Overview}.
\url{http://arxiv.org/abs/1911.11726}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Turak2017Using}{}}%
Turak, E., Brazill-Boast, J., Cooney, T., Drielsma, M., DelaCruz, J.,
Dunkerley, G., Fernandez, M., Ferrier, S., Gill, M., Jones, H., Koen,
T., Leys, J., McGeoch, M., Mihoub, J.-B., Scanes, P., Schmeller, D., \&
Williams, K. (2017). Using the essential biodiversity variables
framework to measure biodiversity change at national scale.
\emph{Biological Conservation}, \emph{213}, 264--271.
\url{https://doi.org/10.1016/j.biocon.2016.08.019}
